---
title: "Final Project"
author: "Kivalina Grove"
date: "2/10/2019"
output: html_document
runtime: shiny
---

## Peer review from Teresa

It's an impressive work! You've gone above and beyond what was required. Your codes will certainly be my refereces if in the future I need to code with leaflet, shiny, tidytext and ggwordcloud. 

### Strength

+ I like how you can use annotation effciently, where you calculate median and paste M in the same line of code. 
+ Although you had a total of 12 complicated datasets to start with, you managed to tidy all dataset with mastery of coding.
+ The biggist strength of your work is to go beyound what was taught in class and select a proper packages to display data visualization. I really admire your motivation in active learning.

### Something I'll change if I were you

+ I am pretty sure this is a personal preference. For me, I like to create as less items as possible in my environment so I can easily access them and I won't get names mess up easily.

### Things I learned

+ The packages you use essentially list out most things I've learned from your work! Really nice work, Kivalina!

```{r installpackages, include=FALSE, eval=FALSE}

install.packages("maptools")
install.packages("leaflet")
install.packages("shiny")
install.packages("rtweet")
install.packages("tidytext")
install.packages("ggwordcloud")

# From a prospective of reproducible document, I'd suggest include #install.packages() or at least list all the required packages up front. I actually got this feedback from one of my reviwer. It makes sense to my so I want to share the same thing with you :)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(dplyr)
library(tidyverse)
library(janitor)
library(lubridate)
```

## Final Project (rough draft stage)

Import Data
```{r importdata}
d1 <- import("./data/201801_hubway_tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "january")
d2 <- import("./data/201802_hubway_tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "february")
d3 <- import("./data/201803_hubway_tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "march")
d4 <- import("./data/201804-hubway-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "april")
d5 <- import("./data/201805-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "may")
d6 <- import("./data/201806-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "june")
d7 <- import("./data/201807-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "july")
d8 <- import("./data/201808-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "august")
d9 <- import("./data/201809-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "september")
d10 <- import("./data/201810-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "october")
d11 <- import("./data/201811-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "november")
d12 <- import("./data/201812-bluebikes-tripdata.csv", setclass = "tbl_df") %>%
  clean_names() %>%
  mutate(month = "december")

stations <- import("./data/Hubway_Stations_as_of_July_2017.csv", setclass = "tbl_df") %>%
  clean_names()

d1 <- d1 %>%
  mutate(birth_year = as.integer(birth_year))

d2 <- d2 %>%
  mutate(birth_year = as.integer(birth_year))

d3 <- d3 %>%
  mutate(birth_year = as.integer(birth_year))

d4 <- d4 %>%
  mutate(birth_year = as.integer(birth_year))

trips <- bind_rows(d1, d2, d3, d4, d5, d6, d7, d8, d9, d10, d11, d12)


# -------- The code below can import multiple datasets. -----------------------------  

d4to12 <- map_df(fs::dir_ls("data", regexp = "-tripdata\\.csv"), 
                 import, setclass = "tbl_df",
                 .id = "dataset") %>% 
  clean_names()


d1to3 <- map_df(fs::dir_ls("data", regexp = "_tripdata\\.csv"), 
                import, setclass = "tbl_df",
                .id = "dataset") %>% 
  clean_names() %>% 
  mutate(birth_year = as.integer(birth_year))

d1tod12 <- bind_rows(d1to3, d4to12)  # This is equivalent to your trips datasest. 

# "Month" is not included though...I don't know why...
# I bet you could extract month info from "startdate" as a work-around method


# I ran in a problem when I tried to import all datasets at once. My guess is not all datasets share the same variable types. "birth_year" is char in d1 to d3, and int in d4 to d12. I had to import datasets seperately. I'm guessing there might be a way to work around it? 

# ----------- The code above can import multiple datasets. ----------------------------

trips_tidy <- trips %>%
  separate(starttime, c("startdate", "starttime"), sep = " ") %>%
  separate(stoptime, c("stopdate", "stoptime"), sep = " ") %>%
  mutate(startdate = ymd(startdate), 
         stopdate = ymd(stopdate),
         weekday = weekdays(startdate)) %>%  
  # It seems neater to me to have 1 `mutate` rather than doing multiple ` mutate` separatly, but I think it's just a personal preference.
  
  #mutate(starttime = hms(starttime), stoptime = hms(stoptime)) %>%
  separate(starttime, c("hour", "minute", "second"))

#filter by:
ridesubset <- trips_tidy %>%
  filter(month == "september", 
         weekday == "Monday", 
         hour == "11", 
         usertype == "Subscriber")

mapsubset <- ridesubset %>%
  mutate(id = seq.int(nrow(ridesubset))) %>%
  select(id,
         start_station_latitude, start_station_longitude, 
         end_station_latitude, end_station_longitude) %>%
  gather(key = group, value = location, -id) %>%
  separate(group, c("group", NA , "where")) %>%
  # Noting really wrong here, I just combined two lines of code into one.
  
  spread(where, location)

```
###### extra code, unused at this point ########
test <- head(d1, 100) %>%
  select(starttime, start_station_latitude, start_station_longitude, end_station_latitude, end_station_longitude) %>%
  gather(key = group, value = location, -starttime) %>%
  separate(group, c("group", "disc", "where")) %>%
  select(-"disc") %>%
  spread(where, location)

test1 <- d1 %>%
  group_by(start_station_latitude, 
           start_station_longitude, 
           end_station_latitude, 
           end_station_longitude) %>%
  count() %>%
  ungroup() %>%
  mutate(n = as.numeric(n)) %>%
  filter(n > 20) %>%
  mutate(id = seq.int(nrow(test1))) %>%
  gather(key = group, value = location, -id, -n) %>%
  separate(group, c("group", "disc", "where")) %>%
  select(-"disc") %>%
  spread(where, location)


d1_filter <- d1 %>%
  separate(starttime, c("date", "time"), sep = " ") %>%
  separate(stoptime, c("date", "time"), sep = " ")
  

Points to Line Function
```{r points_to_line, message=FALSE, warning=FALSE}
library(sp)
library(maptools)

points_to_line <- function(data, long, lat, id_field = NULL, sort_field = NULL) {

  # Convert to SpatialPointsDataFrame
  coordinates(data) <- c(long, lat)

  # If there is a sort field...
  if (!is.null(sort_field)) {
    if (!is.null(id_field)) {
      data <- data[order(data[[id_field]], data[[sort_field]]), ]
    } else {
      data <- data[order(data[[sort_field]]), ]
    }
  }

  # If there is only one path...
  if (is.null(id_field)) {

    lines <- SpatialLines(list(Lines(list(Line(data)), "id")))

    return(lines)

    # Now, if we have multiple lines...
  } else if (!is.null(id_field)) {  

    # Split into a list by ID field
    paths <- sp::split(data, data[[id_field]])

    sp_lines <- SpatialLines(list(Lines(list(Line(paths[[1]])), "line1")))

    # I like for loops, what can I say...
    for (p in 2:length(paths)) {
      id <- paste0("line", as.character(p))
      l <- SpatialLines(list(Lines(list(Line(paths[[p]])), id)))
      sp_lines <- spRbind(sp_lines, l)
    }

    return(sp_lines)
  }
}

# The function code here is way beyond my current knowlege of R, so I can't give any feedback on it. I'll certainly use your code as a reference if I need a similar function!

```


y <- points_to_line(mapsubset, "longitude", "latitude", "id")

stations %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addPolylines(data = y, opacity = 0.1, weight = 3, color = "blue") %>%
    addCircleMarkers(popup = stations$station, 
                   radius = 1, 
                   color = "black")
                   
Visualization 1: map of one bike's travels over the period of a year (here, may to december) 
Use: advertising - maybe markup after to add like "August 8th - Joe and Sam went on a first date"?
```{r mapOneBike}
library(leaflet)
library(shiny)
  
  
popularbike <- trips_tidy %>%
  count(bikeid, sort = TRUE) 
  # count() calls group_by() before and ungroup() after and has a sorting function.

ridesubset <- trips_tidy %>%
  filter(bikeid == head(popularbike$bikeid, 1))

mapsubset <- ridesubset %>%
    mutate(id = seq.int(nrow(ridesubset))) %>%
    select(id,month,
           start_station_latitude, start_station_longitude, 
           end_station_latitude, end_station_longitude) %>%
  gather(key = group, value = location, -id, -month) %>%
  separate(group, c("group", NA, "where")) %>%
  spread(where, location)

y <- points_to_line(mapsubset, "longitude", "latitude", "id")

stations %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addPolylines(data = y, opacity = 0.1, weight = 3, color = "blue") %>%
  addCircleMarkers(popup = stations$station, 
                   radius = 1, 
                   color = "black")

# leaflet is something I think about learning but never get enough motivation to do it.
# After seeing this data visulization, I certainly will put leaflet to my to-explore list.
# This plot is created to show a biker's travel for a period of time. It seems to me the data can only show the traveling displacement (straight line from beginning and end) but not traveling distance/route. Any title of subtitle could help understand the true menssages.
```
 data = test, 
               lng = ~longitude, 
               lat = ~latitude, 
               group = ~group,
               stroke = TRUE,
               opacity = .01,
               color = ~pal(latitude))


??points_to_line()


 addPolylines(data = test1, 
               lat = ~c(test1$start_station_latitude, test1$end_station_latitude), 
               lng = ~c(test1$start_station_longitude, test1$end_station_longitude),
               color = ~pal(n))

 pal <- colorNumeric(
  palette = "BuPu",
  domain = NULL, 
  na.color = NA)
  
ridesubset <- trips_tidy %>%
  filter(month == input$month, 
         weekday == input$weekday, 
         hour == input$hour, 
         usertype == "Subscriber")

  mapsubset <- ridesubset %>%
    mutate(id = seq.int(nrow(ridesubset))) %>%
    select(id,
         start_station_latitude, start_station_longitude, 
         end_station_latitude, end_station_longitude) %>%
  gather(key = group, value = location, -id) %>%
  separate(group, c("group", "disc", "where")) %>%
  select(-"disc") %>%
  spread(where, location)

y <- points_to_line(mapsubset, "longitude", "latitude", "id")

stations %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addPolylines(data = y, opacity = 0.1, weight = 3, color = "blue") %>%
    addCircleMarkers(popup = stations$station, 
                   radius = 1, 
                   color = "black")

```{r mapcolorbyStationStart}
start_count <- trips_tidy %>%
  count(start_station_latitude, start_station_longitude) 


 pal <- colorNumeric(
  palette = "inferno",
  domain = NULL, 
  na.color = NA)
 
stations %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(popup = ~paste0(stations$station, "</br>", "Trips Started: ", start_count$n, "</br>", "Number of Docks: ", stations$number_of_docks),
                   radius = stations$number_of_docks/3,
                   color = ~pal(start_count$n),
                   fillOpacity = .6,
                   weight = 5,
                   fill = TRUE,
                   stroke = FALSE) %>%
  addLegend(title = "Rides Started", pal = pal, values = c(0:53846), position = "bottomright")

# I am guessing this data visulization is about how popluar the station is when renting the bikes? Any title or subtitle may help deliver the menssages. 
```

```{r mapcolorbyStationEnd}
end_count <- trips_tidy %>%
  count(end_station_latitude, end_station_longitude)

 pal <- colorNumeric(
  palette = "inferno",
  domain = NULL, 
  na.color = NA)
 
(stations %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(popup = ~paste0(stations$station, "</br>", "Rides Completed: ", end_count$n, "</br>", "Number of Docks: ", stations$number_of_docks),
                   radius = stations$number_of_docks/3,
                   color = ~pal(start_count$n),
                   fillOpacity = .6,
                   weight = 5,
                   fill = TRUE,
                   stroke = FALSE) %>%
  addLegend(title = "Rides Completed", pal = pal, values = c(0:50000), position = "bottomright"))
 
 # I am not familar with leaflet() so I am curious why you need ( and ) at the beginning and end? 
```
To do: set default map views, snap back to view, figure out how to make legend non-transparent?
- search among stations?
- size by number of bikes available?

Visual #3: Histogram of bike rides in the month of September
```{r histogram}
trips_september <- trips_tidy %>%
  filter(month == "september", tripduration < 3600) %>%
  mutate(tripminutes = (tripduration/60))

(plot <- trips_september %>%
  ggplot(aes(x = tripminutes)) +
  geom_density(fill = "black", 
               color = "white", 
               alpha = 0.7) +
  geom_vline(xintercept = median(trips_september$tripminutes), 
             color = "black", 
             linetype = "dashed") +
  annotate("text", 
           label = paste("M =", round(median(trips_september$tripminutes), 2)), 
           y = 0.07, x = 6, 
           color = "black", 
           size = 3) +
  geom_density(data = filter(trips_september, weekday == "Saturday"), 
               fill = "#335DFF", 
               color = "white", 
               alpha = 0.5) +
  geom_vline(xintercept = median(filter(trips_september, weekday == "Saturday")$tripminutes),
             color = "#335DFF", 
             linetype = "dashed") +
  annotate("text", 
           label = paste("M =", round(median(filter(trips_september, 
                                                    weekday == "Saturday")$tripminutes), 2)),
           y = 0.073, x = 6, 
           color = "#335DFF", 
           size = 3) +
    labs(x = "Trip Duration in Minutes", y = "Density", 
         title = "Average Ride Length in September by Weekday",
         caption = "where M represents the median value of each distribution", size = 5) +
    theme_minimal() +
    annotate("text", label = c("September\n   Overall", "Saturday"),
             y = c(0.045, 0.025), x = c(15, 22), 
             hjust = 0, size = 3, color = c("black", "blue")))

weekday_list <- as.list(unique(trips_september$weekday))
```

```{r}
inputPanel(
  selectInput("weekday", label = "Weekday",
              choices = weekday_list,
              selected = 1)
)

# The use of shiny package is again beyond my knowlege so apologize for not being able to comment anything. 
```

```{r shinyPlot, echo = FALSE}
renderPlot({
  (plot <- trips_september %>%
  ggplot(aes(x = tripminutes)) +
  geom_density(fill = "black", 
               color = "white",
               alpha = 0.7) +
  geom_vline(xintercept = median(trips_september$tripminutes), 
             color = "black", 
             linetype = "dashed") +
  annotate("text", 
           label = paste("M =", round(median(trips_september$tripminutes), 2)), 
           y = 0.073, x = 6, 
           color = "black", size = 5) +
  geom_density(data = filter(trips_september, weekday == input$weekday), 
               fill = "#335DFF", 
               color = "white", 
               alpha = 0.5) +
  geom_vline(xintercept = median(filter(trips_september, 
                                        weekday == input$weekday)$tripminutes),
             color = "#335DFF", linetype = "dashed") +
  annotate("text", 
           label = paste("M =", round(median(filter(trips_september, 
                                                    weekday == input$weekday)$tripminutes),2)), 
           y = 0.076, x = 6, 
           color = "#335DFF", size = 5) +
    labs(x = "Trip Duration in Minutes", y = "Density", 
         title = "September Average Ride Length by Weekday",
         caption = "where M represents the median value of each distribution") +
    theme_minimal(base_size = 15) +
    annotate("text", label = c("September\n   Overall", paste(input$weekday)),
             y = c(0.045, 0.025), x = c(15, 22), 
             hjust = 0, size = 5, color = c("black", "blue")))
}, width = 600, height = 600)
```

previous code:

trips_season <- trips_tidy %>%
  mutate(season = ifelse(month == "december" | month == "january" | month == "february", 
                         yes = "winter", 
                         no = ifelse(month == "march" | month == "april" | month == "may", 
                                     yes = "spring", 
                                     no = ifelse(month == "june" | month == "july" | month == "august",
                                                 yes = "summer", 
                                                 no = "fall")))) %>%
  filter(tripduration < 3600)
  

Visual 2: wordcloud and accompanying bar chart for word frequency
```{r wordcloud}
library(rtweet)
library(tidytext)
library(ggwordcloud)
#twitterdata <- search_tweets("boston bike OR boston bicycle OR bluebikes OR bluebikesboston OR #BBWinterChallenge OR ridebluebikes", n = 18000, include_rts = FALSE)
#twitterdata2 <- search_tweets("bikeshare", n = 18000, include_rts = FALSE)

#save(twitterdata2, file = "tweets.RData")

tweets <- import("tweets.RData")

tweets_tidy <- tweets %>%
  select(screen_name, created_at, text, hashtags) %>%
  unnest_tokens(word, text, token = "words")

c_stop_words <- bind_rows(data_frame(word = c("t.co", "https", "i’m", "и", 
                                              "10", "de", "bikeshare"), 
                                          lexicon = c("custom", "custom", "custom", "custom", 
                                                      "custom", "custom", "custom")), 
                                          stop_words)
tweets_tidy <- tweets_tidy %>%
  anti_join(c_stop_words)

tweets_count <- tweets_tidy %>%
  count(word, sort = TRUE)

tweets_count %>%
  filter(n > 15) %>%
  mutate(word = reorder(as.factor(word), n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(alpha = 0.7, fill = "#4286f4", color = "white") +
  coord_flip() +
  theme_light() +
  theme(text = element_text(size = 20))

# Is there any reason you create tweets_count separately? or is it ok to comebine two chunks of code into one?

tweets_tidy <- tweets_tidy %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE)

tweets_positive <- tweets_tidy %>%
  filter(sentiment == "positive") %>%
  head(15) %>%
  mutate(word = reorder(as.factor(word), n))

tweets_negative <- tweets_tidy %>%
  filter(sentiment == "negative") %>%
  head(15) %>%
  mutate(word = reorder(as.factor(word), n))

ggplot(tweets_positive,aes(x = word, y = n)) +
  geom_col(alpha = 0.7, fill = "#4286f4", color = "white") +
  coord_flip() +
  theme_light() +
  theme(text = element_text(size = 20)) +
  labs(x = NULL, y = NULL, title = "Top 15 Positively Valenced Words")

ggplot(tweets_negative,aes(x = word, y = n)) +
  geom_col(alpha = 0.7, fill = "#f44271", color = "white") +
  coord_flip() +
  theme_light() +
  theme(text = element_text(size = 20)) +
  labs(x = NULL, y = NULL, title = "Top 15 Negatively Valenced Words")

set.seed(15)

tweets_angled <- tweets_tidy %>%
  group_by(sentiment) %>%
  top_n(15) %>%
  mutate(angle = 90 * sample(c(0, 1), n(), 
                             replace = TRUE, 
                             prob = c(50, 50)))

(sentcloud <- ggplot(tweets_angled, aes(label = word, size = n, 
                                        color = sentiment, angle = angle)) +
  geom_text_wordcloud_area(rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  theme_minimal())

# why do you need ( and ) at the beginning and end? maybe it's something with ggcloudwords?
```

